Supporting Information : Asymmetric tail-association in extreme climatic events affects population-extinction risk

### 2019-10-11 SuppMat.Rmd
started editing intro...continued

---
title: "Supporting Information : Asymmetric tail-dependence in extreme climatic events affects population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

# flip Clayton copula 
mtc_r<-1-mtc_l

pdf("./Results/fig_4_copula.pdf", height=3, width=12)
op<-par(mfrow=c(1,4),mar=c(4,4,2,2),mgp=c(2,1,0))

# extreme left tail
plot(xtc_l[,1],xtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
legend(x="topleft",legend="A",bty="n",cex=2)

# moderate left tail
plot(mtc_l[,1],mtc_l[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
legend(x="topleft",legend="B",bty="n",cex=2)

# moderate right tail
plot(mtc_r[,1],mtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
legend(x="topleft",legend="C",bty="n",cex=2)

# extreme right tail
plot(xtc_r[,1],xtc_r[,2],col=rgb(0,0,0,0.05),pch=19,xlab="u",ylab="v",cex.lab=2)
legend(x="topleft",legend="D",bty="n",cex=2)

@@

<!--FIG:  age structure model results for moderate tail dep. (C, SC)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) dependence.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations. Fecundity of adults and survival of eggs have moderate tail dependence in either left or right tail in their corresponding copula structure. Moderate left tail dependence (or right tail dependence) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
\end{center}
\end{figure}

<!--FIG:  summary: all nonlinear model results for local dispersion, moderate noise : extrisk vs. d-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_moderatenoise.pdf}
\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
\end{center}
\end{figure}

@@

<!--FIG:  summary: all nonlinear model results for global dispersion, moderate noise : extrisk vs. d-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_moderatenoise.pdf}
\caption[Analogous figure as that of Fig. \ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.]{Analogous figure as that of Fig. \ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail dependence in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
\end{center}
\end{figure}

@@

<!--FIG: Ricker model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_ricker_risk_local_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG : Hassell model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_hassell_risk_local_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

@@

<!--FIG: Ricker model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_ricker_risk_global_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG : Hassell model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_hassell_risk_global_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_msmith_risk_global_disp_scl_0.8_mtd}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG : all nonlinear model results with moderate noise, d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_moderatenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_moderatenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_moderatenoise.pdf}\\
\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
tail dependent noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
tail dependent noises; Moderate left tail dependence (or right tail dependence) was generated 
by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between
any two out of $n=5$ patches.\label{fig_extrisk_d0_mtd}}
\end{center}
\end{figure*}

# Generating matrix with extreme tail dependence  \label{extr_tail_dep}

For extreme tail-dependence algorithm, a $N \times n$ matrix, M was generated of independent 
standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->
Then, each row of M was identified for a first set of modifications
randomly and independently with a 50\% chance; for those rows identified, every
element of the row was replaced with the absolute
value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of 
its own absolute value. This produces right-tail dependent
matrix, normally distributed for each column. If left-tail 
dependent noise is desired, take the negative of the entire matrix.
Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
generating extreme tail dependence matrix is flexible enough, so that 
matrix M can be generated initially with independent random draws from a normal 
distribution with a specified mean and standard deviation.

>>>>

---
title: "Supporting Information : Asymmetric tail-association in extreme climatic events affects population-extinction risk"
author: "Shyamolina Ghosh, Daniel Reuman"
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

# flip Clayton copula 
mtc_r<-1-mtc_l
```

```{r plot_intro_fig, echo=F, results="hide"}

# for plotting purpose I just showed only first 500 points
pdf("./Results/introfig_copula.pdf", height=3, width=12)
op<-par(mfrow=c(1,4),mar=c(6,6,2,2),mgp=c(3,1,0))

# extreme left tail
plot(xtc_l[1:500,1],xtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
legend(x="topleft",legend="A",bty="n",cex=2.5)

# moderate left tail
plot(mtc_l[1:500,1],mtc_l[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
legend(x="topleft",legend="B",bty="n",cex=2.5)

# moderate right tail
plot(mtc_r[1:500,1],mtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
legend(x="topleft",legend="C",bty="n",cex=2.5)

# extreme right tail
plot(xtc_r[1:500,1],xtc_r[1:500,2],col=rgb(0,0,0,0.1),pch=19,xlab="u",ylab="v",cex.lab=2.5,cex.axis=2)
legend(x="topleft",legend="D",bty="n",cex=2.5)

saveRDS(ktau_common,"./Results/ktau_common.RDS")
```

@@

<!--FIG:  age structure model results for moderate tail dep. (C, SC)-->
\begin{figure}[!h]
\begin{center}
\textbf{ \hspace{0.5 cm} (A) \hspace{4.4 cm} (B)} \\
\vspace{-0.2 cm}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_linear.pdf}
\includegraphics[width=5 cm]{./Results/age_str_results/extrisk_age_str_C_SC_nonlinear_omg_100.pdf}
\caption[Extinction risk from single patch age structured models with time when vital rates have either left (C) or right tail (SC) association.]{Extinction risk (shading represents 95\% confidence interval) from single patch age structured models with time for 10000 simulations. Fecundity of adults and survival of eggs have moderate tail association in either left or right tail in their corresponding copula structure. Moderate left tail association (or right tail association) was generated by a bivariate Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6, $(A)$ for linear model ($\Omega = \infty$) and for $(B)$ non-linear model ($\Omega = 100$).\label{fig_extrisk_agestr_mtd}}
\end{center}
\end{figure}

<!--FIG:  summary: all nonlinear model results for local dispersion, moderate noise : extrisk vs. d-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_local_disp_moderatenoise.pdf}
\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_allmodel_localdisp_etd} in the maintext but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_localdisp_mtd}}
\end{center}
\end{figure}

@@

<!--FIG:  summary: all nonlinear model results for global dispersion, moderate noise : extrisk vs. d-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/allmodel_ext_risk_global_disp_moderatenoise.pdf}
\caption[Analogous figure as that of Fig. \ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.]{Analogous figure as that of Fig. \ref{fig_extrisk_allmodel_globaldisp_mtd} but with moderate tail association in spatially correlated noises.\label{fig_extrisk_allmodel_globaldisp_mtd}}
\end{center}
\end{figure}

@@

<!--FIG: Ricker model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_ricker_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_ricker_risk_local_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG : Hassell model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_hassell_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_hassell_risk_local_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_msmith_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_msmith_risk_local_disp_scl_0.8_mtd}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_pennycuick_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_pennycuick_r_vary_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_verhulst_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_verhulst_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, local dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_local_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with local dispersal.]{Analogous figure as of Fig. \ref{fig_abrewer_risk_local_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_local_disp_abrewer_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

@@

<!--FIG: Ricker model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/ricker_scl_1/ricker_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Ricker metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_ricker_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_ricker_risk_global_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG : Hassell model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/hassell_scl_1/hassell_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Hassell metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_hassell_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_hassell_risk_global_disp_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Maynard smith model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/msmith_scl_0.8/msmith_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Maynard Smith metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_msmith_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_msmith_risk_global_disp_scl_0.8_mtd}}
\end{center}
\end{figure}

<!--FIG: Pennycuick model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/pennycuick_scl_1/pennycuick_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Pennycuick metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_pennycuick_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_pennycuick_r_vary_scl_1_mtd}}
\end{center}
\end{figure}

<!--FIG: Verhulst model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/verhulst_scl_0.6/verhulst_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Verhulst metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_verhulst_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_verhulst_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG: Austin-Brewer model, global dispersal, moderate noise-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=12 cm]{./Results/austinbrewer_scl_0.6/abrewer_ext_risk_global_disp_moderatenoise.pdf}\\
\caption[Effect of moderate tail-dependent environmental noises on extinction risk for Austin-Brewer metapopulation model with global dispersal.]{Analogous figure as of Fig. \ref{fig_abrewer_risk_global_disp_scl_1_etd} but with moderate left- and right-tail dependent noises; Moderate left tail association (or right tail association) was generated by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between any two out of $n=5$ patches. \label{fig_risk_global_disp_abrewer_r_vary_scl_0.6_mtd}}
\end{center}
\end{figure}

<!--FIG : all nonlinear model results with moderate noise, d=0, extrisk vs. r-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=10 cm]{./Results/common_legend_moderate_noise.pdf}\\
\includegraphics[width=5 cm]{./Results/ricker_scl_1/ricker_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/hassell_scl_1/hassell_extrisk_d0_moderatenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/msmith_scl_0.8/msmith_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/pennycuick_scl_1/pennycuick_extrisk_d0_moderatenoise.pdf}\\
\vspace{-0.6 cm}
\includegraphics[width=5 cm]{./Results/verhulst_scl_0.6/verhulst_extrisk_d0_moderatenoise.pdf}
\includegraphics[width=5 cm]{./Results/austinbrewer_scl_0.6/abrewer_extrisk_d0_moderatenoise.pdf}\\
\caption[Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
tail dependent noises.]{Analogous figure as that of Fig. \ref{MT-fig_extrisk_d0_etd} but with moderarte 
tail dependent noises; Moderate left tail association (or right tail association) was generated 
by Clayton (or Survival Clayton) copula with Kendall correlation coefficient = 0.6 between
any two out of $n=5$ patches.\label{fig_extrisk_d0_mtd}}
\end{center}
\end{figure*}

# Generating matrix with extreme tail association  \label{extr_tail_dep}

For extreme tail-association algorithm, a $N \times n$ matrix, M was generated of independent 
standard-normal random draws (in general). <!--where $N = numsteps(=25) \times numsims(=10000)$ : only for 2d case-->
Then, each row of M was identified for a first set of modifications
randomly and independently with a 50\% chance; for those rows identified, every
element of the row was replaced with the absolute
value of the first element of the row (including the first element itself).
For other rows, every element in the row was replaced by the negative of 
its own absolute value. This produces right-tail dependent
matrix, normally distributed for each column. If left-tail 
dependent noise is desired, take the negative of the entire matrix.
Our [code](https://github.com/sghosh89/ERC/blob/master/ExtremeTailDep.R) for
generating extreme tail association matrix is flexible enough, so that 
matrix M can be generated initially with independent random draws from a normal 
distribution with a specified mean and standard deviation.
